---
title: "Final project"
output: html_document
---

###Rubric

x pts | Define research parameters of question - exploring magnitudes of integration and potential evolvability

x pts | Pull in tree data  

x pts | Simulate data for magnitudes of integration following Grabowski and Porto 2017

x pts | Create visual interpretation 

x pts | run a power analysis based on these results to determine minimum sample size needed for my dissertation data collection


####

Important Definitions:
- Magnitude of integration = The average of squared correlations among traits
- Mean integration = Thee average relative degree to which evolvability is reduced due to conditioning on other traits
- Mean respondability = The average length of the predicted response to selection; how quickly a population can respond to selection
- Mean flexibility = The average cosine of angle between the direction of selection and the response vector
- Mean evolvability = The average length of the multivariate response in the direction of selection for a given phenotypic variance/covariance matrix
- Mean conditional evolvability = The average length of the multivariate response in the direction of selection for given phenotypic variance/covariance matrix




Load tree data 

``` {r}
library(rotl)
library(ape)
taxa <- rotl::tnrs_match_names(name=c("Mustelidae", "Procyonidae", "Mephitidae", "Phocidae", "Otariidae", "Ursidae", "Canidae", "Herpestidae", "Eupleridae", "Hyaenidae", "Felidae"))
tree <- rotl::tol_subtree(ott_id(taxa))
plot(tree, cex = 0.8, label.offset = 0.1)

```


Simulate integration magnitude data for species in tree

``` {r}
library(clusterGeneration)
library(truncnorm)
library(matrixcalc)
library(foreach)
library(doParallel)
library(moments)

```


Power analysis - effect size, sample size, significance level, power

```{r}
library(pwr)



```


